---
title: "cryptocurrencies"
author: "JiaJun Neo"
date: "19/04/2018"
output:
  html_document :                   # custom style for HTML
    # theme of html document
    # theme of code highlight                                 
    # table of contents
    theme       : paper             # "default", "cerulean", "journal",
                                    # "flatly", "readable", "spacelab",
                                    # "united", "cosmo", "lumen", "paper", 
                                    # "sandstone", "simplex", "yeti"
    highlight   : pygments          # "default", "tango", "pygments",
                                    # "kate",  "monochrome", "espresso",
                                    # "zenburn", "haddock", "textmate"
    toc         : true
    toc_depth   : 2
    
  pdf_document  : default           # default style for PDF
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 12, fig.height = 4)
```
***********************************************************

# Initial Setup

Install and load the essential R packages for Time Series analysis -- `tseries` and `forecast`.

```{r}
# install.packages("tseries")
# install.packages("forecast")
# install.packages("ggplot2")
library(tseries)
library(forecast)
library("ggplot2")
```

Now we read and view the structure of the csv file downloaded from Kaggle.com 

```{r}
all_crypto <- read.csv("crypto-markets.csv")

# Structure of the CSV File
str(all_crypto)
```

# Preprocessing

The data is in the format where all the cryptocurrencies are in the symbols and is actually stacked into a row.

This section will be focusing on getting the data into what we actually want.

The cryptocurrencies that we want are:

Bitcoin, Ethereum, Ripple, Bitcoin Cash and Litecoin

The columns that we need:

Symbol, Date, Open and Close

The date of cryptocurrencies that we need: 
"2017-04-01" - "2018-04-15"

```{r}
# ==============================================================
# Pre-processing
# ==============================================================

# Get the cryptocurrencies that we want
# ==============================================================
# BTC, ETH, XRP, BCH, LTC
# ==============================================================

btc <- all_crypto[all_crypto$symbol == 'BTC',]
eth <- all_crypto[all_crypto$symbol == 'ETH',]
xrp <- all_crypto[all_crypto$symbol == 'XRP',]
bch <- all_crypto[all_crypto$symbol == 'BCH',]

ltc <- all_crypto[all_crypto$symbol == 'LTC',]

# Filter the columns:
# ===============================================================
# Column we need : symbol, date, open, close 
# ===============================================================

btc_final <- data.frame(btc$symbol, btc$date, btc$open, btc$close)
colnames(btc_final) <- c("symbol", "date", "open", "close")

eth_final <- data.frame(eth$symbol, eth$date, eth$open, eth$close)
colnames(eth_final) <- c("symbol", "date", "open", "close")

xrp_final <- data.frame(xrp$symbol, xrp$date, xrp$open, xrp$close)
colnames(xrp_final) <- c("symbol", "date", "open", "close")

bch_final <- data.frame(bch$symbol, bch$date, bch$open, bch$close)
colnames(bch_final) <- c("symbol", "date", "open", "close")

ltc_final <- data.frame(ltc$symbol, ltc$date, ltc$open, ltc$close)
colnames(ltc_final) <- c("symbol", "date", "open", "close")

# Format the date & symbol
# ================================================================
# Current implementation is Factor
# ================================================================
btc_final$date <- as.Date(as.factor(btc_final$date))
btc_final$symbol <- as.character(as.factor(btc_final$symbol))

eth_final$date <- as.Date(as.factor(eth_final$date))
eth_final$symbol <- as.character(as.factor(eth_final$symbol))

xrp_final$date <- as.Date(as.factor(xrp_final$date))
xrp_final$symbol <- as.character(as.factor(xrp_final$symbol))

bch_final$date <- as.Date(as.factor(bch_final$date))
bch_final$symbol <- as.character(as.factor(bch_final$symbol))

ltc_final$date <- as.Date(as.factor(ltc_final$date))
ltc_final$symbol <- as.character(as.factor(ltc_final$symbol))

# ==============================================================
# Sorting the date 
# ==============================================================

btc_sorted_final <- btc_final[btc_final$date >= "2017-04-01" & btc_final$date <= "2018-04-15",]
eth_sorted_final <- eth_final[eth_final$date >= "2017-04-01" & eth_final$date <= "2018-04-15",]
xrp_sorted_final <- xrp_final[xrp_final$date >= "2017-04-01" & xrp_final$date <= "2018-04-15",]
bch_sorted_final <- bch_final[bch_final$date >= "2017-04-01" & bch_final$date <= "2018-04-15",]
ltc_sorted_final <- ltc_final[ltc_final$date >= "2017-04-01" & ltc_final$date <= "2018-04-15",]

```

# BITCOIN

Bitcoin is currently the number 1 cryptocurrency in the world. This is due to the first mover advantage and most people heard of bitcoin. 

For all cryptocurrencies, they showed a huge spike in December 2017 when Bitcoin actually beat a lot of company in value which was personally absurd to me. 

```{r }
# plot open price
ggplot() + 
  geom_line(data = btc_sorted_final, aes(x = date, y = open, color="Open Price")) +
  geom_line(data = btc_sorted_final, aes(x = date, y = close, color="Close Price"))

summary(btc_sorted_final$open)
summary(btc_sorted_final$close)
```

### Moving Average of Bitcoin [Weekly]

Since the price of cryptocurrencies are so volatile. I decided to use a rolling moving average to stabilize the data slightly.

```{r }
btc_latest_model <- btc_sorted_final[btc_sorted_final$date >= "2017-04-15" & btc_sorted_final$date <= "2018-04-15",]
btc_latest_model$open_ma = ma(btc_latest_model$open, order=7) # Weekly Open Moving Average

btc_latest_model$close_ma = ma(btc_latest_model$close, order=7) # Weekly Close Moving Average

ggplot() + 
  geom_line(data = btc_latest_model, aes(x = date, y = open, color="open prices")) +
  geom_line(data = btc_latest_model, aes(x = date, y = close, color="close prices")) +
  geom_line(data = btc_latest_model, aes(x = date, y = open_ma, color="weekly open ma")) + 
  geom_line(data = btc_latest_model, aes(x = date, y= close_ma, color="weekly close ma"))

```

### Creating Stationary Data for Open Moving Average 

```{r}
btc_open_ma <- ts(na.omit(btc_latest_model$open_ma))

# Stationarity
adf.test(btc_open_ma, alternative = "stationary")
# Stop if p-value < 0.05

```

```{r }
btcLog <- log(btc_open_ma)
plot(btcLog, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(btcLog)
# Stop if p-value < 0.05
```

```{r }
btcLogDiff <- diff(btcLog)
plot(btcLogDiff, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(btcLogDiff)
# Stop if p-value < 0.05
```

### ACF and PACF of Bitcoin Open Moving Average

```{r }
op <- par(mfrow = c(1, 2))
acf(btcLogDiff, main='ACF For BTC Differenced Series')
pacf(btcLogDiff, main='PACF For BTC Differenced Series')
par(op)
```

### Arima of Open Price Based on Guess

```{r }
arimaFit <- arima(btcLogDiff, order = c(1,0,0))
arimaFit            # check the coefficients
plot(arimaFit)


arimaFitFC <- forecast(btcLogDiff, model = arimaFit, h = 10)
plot(btcLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaFitFC$fitted, type="l", col = "red", lwd = 2)

plot(arimaFitFC)
```

```{r }
tsdisplay(residuals(arimaFit), lag.max = 80)
```

### Finding Optimal Arima of Bitcoin

```{r}
auto.arima(btcLogDiff,                   # the dataset
           seasonal = FALSE,            # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

### Arima Optimal Plots

```{r }
arimaOpt <- arima(btcLogDiff, order = c(1,0,1))
arimaOpt
plot(arimaOpt)

arimaOptFC <- forecast(btcLogDiff, model = arimaOpt, h = 10)
plot(btcLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaOptFC$fitted, type="l", col = "red", lwd = 2)
# Beyond the training dataset
plot(arimaOptFC)
```

### Bitcoin Seasonal? 

```{r}
auto.arima(btcLogDiff,                   # the dataset
           seasonal = TRUE,             # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

No! Bitcoin is not seasonal and thus we just utilize the auto.arima above.

```{r}
# Check the performance/accuracy
tsdisplay(residuals(arimaOpt), lag.max = 80)
```

### Bitcoin Close Price Stationarity 

```{r}
btc_close_ma <- ts(na.omit(btc_latest_model$close_ma))

# Stationarity
adf.test(btc_close_ma, alternative = "stationary")
# Stop if p-value < 0.05
```

```{r }
btcCloseLog <- log(btc_close_ma)
plot(btcCloseLog, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(btcCloseLog)
# Stop if p-value < 0.05
```

```{r }
btcCloseLogDiff <- diff(btcCloseLog)
plot(btcCloseLogDiff, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(btcCloseLogDiff)
# Stop if p-value < 0.05
```

### Bitcoin Close Price's ACF and PACF

```{r }
op <- par(mfrow = c(1, 2))
acf(btcCloseLogDiff, main='ACF For BTC Differenced Series')
pacf(btcCloseLogDiff, main='PACF For BTC Differenced Series')
par(op)
```

### Bitcoin Close Price Arima Guess

```{r }
# Arima Based on guessing
arimaFit <- arima(btcCloseLogDiff, order = c(4,0,2))
arimaFit            # check the coefficients
plot(arimaFit)

arimaFitFC <- forecast(btcCloseLogDiff, model = arimaFit, h = 10)
plot(btcCloseLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaFitFC$fitted, type="l", col = "red", lwd = 2)

plot(arimaFitFC)
```

```{r}
tsdisplay(residuals(arimaFit), lag.max = 80)
```

### Optimal Arima Close Price
```{r}
auto.arima(btcCloseLogDiff,                   # the dataset
           seasonal = FALSE,            # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

### Bitcoin Close Forecast Plot

```{r}
arimaOpt <- arima(btcCloseLogDiff, order = c(1,0,1))
arimaOpt
plot(arimaOpt)

# Forecast using optimal ARIMA
# Within the training dataset
arimaOptFC <- forecast(btcCloseLogDiff, model = arimaOpt, h = 10)
plot(btcCloseLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaOptFC$fitted, type="l", col = "red", lwd = 2)
# Beyond the training dataset
plot(arimaOptFC)
```

```{r}
tsdisplay(residuals(arimaOpt), lag.max = 80)
```

We know moving average bitcoin is not seasonal! Thus we are moving on!

# Ethereum

Ethereum is the 2nd cryptocurrency in the world. It can create smart contracts today the price is at 50% loss compared to Dec 2017.

```{r }
# plot open price
ggplot() + 
  geom_line(data = eth_sorted_final, aes(x = date, y = open, color="Open Price")) +
  geom_line(data = eth_sorted_final, aes(x = date, y = close, color="Close Price"))

summary(eth_sorted_final$open)
summary(eth_sorted_final$close)
```

### Moving Average of Ethereum [Weekly]

Since the price of cryptocurrencies are so volatile. I decided to use a rolling moving average to stabilize the data slightly.

```{r }
eth_latest_model <- eth_sorted_final[eth_sorted_final$date >= "2017-04-15" & eth_sorted_final$date <= "2018-04-15",]
eth_latest_model$open_ma = ma(eth_latest_model$open, order=7) # Weekly Open Moving Average

eth_latest_model$close_ma = ma(eth_latest_model$close, order=7) # Weekly Close Moving Average

ggplot() + 
  geom_line(data = eth_latest_model, aes(x = date, y = open, color="open prices")) +
  geom_line(data = eth_latest_model, aes(x = date, y = close, color="close prices")) +
  geom_line(data = eth_latest_model, aes(x = date, y = open_ma, color="weekly open ma")) + 
  geom_line(data = eth_latest_model, aes(x = date, y= close_ma, color="weekly close ma"))

```

### Creating Stationary Data for Open Moving Average 

```{r}
eth_open_ma <- ts(na.omit(eth_latest_model$open_ma))

# Stationarity
adf.test(eth_open_ma, alternative = "stationary")
# Stop if p-value < 0.05

```

```{r }
ethLog <- log(eth_open_ma)
plot(ethLog, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(ethLog)
# Stop if p-value < 0.05
```

```{r }
ethLogDiff <- diff(ethLog)
plot(ethLogDiff, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(ethLogDiff)
# Stop if p-value < 0.05
```

### ACF and PACF of Bitcoin Open Moving Average

```{r }
op <- par(mfrow = c(1, 2))
acf(ethLogDiff, main='ACF For ETH Differenced Series')
pacf(ethLogDiff, main='PACF For ETH Differenced Series')
par(op)
```

### Arima of Open Price Based on Guess

```{r }
arimaFit <- arima(ethLogDiff, order = c(4,0,5))
arimaFit            # check the coefficients
plot(arimaFit)


arimaFitFC <- forecast(ethLogDiff, model = arimaFit, h = 10)
plot(ethLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaFitFC$fitted, type="l", col = "red", lwd = 2)

plot(arimaFitFC)
```

```{r }
tsdisplay(residuals(arimaFit), lag.max = 80)
```

### Finding Optimal Arima of Ethereum

```{r}
auto.arima(ethLogDiff,                   # the dataset
           seasonal = FALSE,            # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

### Arima Optimal Plots

```{r }
arimaOpt <- arima(ethLogDiff, order = c(3,0,5))
arimaOpt
plot(arimaOpt)

arimaOptFC <- forecast(ethLogDiff, model = arimaOpt, h = 10)
plot(ethLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaOptFC$fitted, type="l", col = "red", lwd = 2)
# Beyond the training dataset
plot(arimaOptFC)
```

### Ethereum Seasonal? 

```{r}
auto.arima(ethLogDiff,                   # the dataset
           seasonal = TRUE,             # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

No! Ethereum is not seasonal and thus we just utilize the auto.arima above.

```{r}
# Check the performance/accuracy
tsdisplay(residuals(arimaOpt), lag.max = 80)
```

### Ethereum Close Price Stationarity 

```{r}
eth_close_ma <- ts(na.omit(eth_latest_model$close_ma))

# Stationarity
adf.test(eth_close_ma, alternative = "stationary")
# Stop if p-value < 0.05
```

```{r}
ethCloseLog <- log(eth_close_ma)
plot(ethCloseLog, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(ethCloseLog)
# Stop if p-value < 0.05
```

```{r}
ethCloseLogDiff <- diff(ethCloseLog)
plot(ethCloseLogDiff, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(ethCloseLogDiff)
# Stop if p-value < 0.05
```

### Ethereum Close Price's ACF and PACF

```{r }
op <- par(mfrow = c(1, 2))
acf(ethCloseLogDiff, main='ACF For BTC Differenced Series')
pacf(ethCloseLogDiff, main='PACF For BTC Differenced Series')
par(op)
```

### Ethereum Close Price Arima Guess

```{r }
# Arima Based on guessing
arimaFit <- arima(ethCloseLogDiff, order = c(4,0,2))
arimaFit            # check the coefficients
plot(arimaFit)

arimaFitFC <- forecast(ethCloseLogDiff, model = arimaFit, h = 10)
plot(ethCloseLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaFitFC$fitted, type="l", col = "red", lwd = 2)

plot(arimaFitFC)
```

```{r}
tsdisplay(residuals(arimaFit), lag.max = 80)
```

### Optimal Arima Close Price
```{r}
auto.arima(ethCloseLogDiff,                   # the dataset
           seasonal = FALSE,            # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

### Ethereum Close Forecast Plot

```{r}
arimaOpt <- arima(ethCloseLogDiff, order = c(1,0,1))
arimaOpt
plot(arimaOpt)

# Forecast using optimal ARIMA
# Within the training dataset
arimaOptFC <- forecast(ethCloseLogDiff, model = arimaOpt, h = 10)
plot(ethCloseLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaOptFC$fitted, type="l", col = "red", lwd = 2)
# Beyond the training dataset
plot(arimaOptFC)
```

```{r}
tsdisplay(residuals(arimaOpt), lag.max = 80)
```

# Ripple

Ripple is the 3rd cryptocurrency in the world. It aim is to help deal with bank transaction easily and faster. $3 was the highest price.

```{r }
# plot open price
ggplot() + 
  geom_line(data = xrp_sorted_final, aes(x = date, y = open, color="Open Price")) +
  geom_line(data = xrp_sorted_final, aes(x = date, y = close, color="Close Price"))

summary(xrp_sorted_final$open)
summary(xrp_sorted_final$close)
```

### Moving Average of Ripple [Weekly]

Since the price of cryptocurrencies are so volatile. I decided to use a rolling moving average to stabilize the data slightly.

```{r }
xrp_latest_model <- xrp_sorted_final[xrp_sorted_final$date >= "2017-04-15" & xrp_sorted_final$date <= "2018-04-15",]
xrp_latest_model$open_ma = ma(xrp_latest_model$open, order=7) # Weekly Open Moving Average

xrp_latest_model$close_ma = ma(xrp_latest_model$close, order=7) # Weekly Close Moving Average

ggplot() + 
  geom_line(data = xrp_latest_model, aes(x = date, y = open, color="open prices")) +
  geom_line(data = xrp_latest_model, aes(x = date, y = close, color="close prices")) +
  geom_line(data = xrp_latest_model, aes(x = date, y = open_ma, color="weekly open ma")) + 
  geom_line(data = xrp_latest_model, aes(x = date, y= close_ma, color="weekly close ma"))

```

### Creating Stationary Data for Open Moving Average 

```{r}
xrp_open_ma <- ts(na.omit(xrp_latest_model$open_ma))

# Stationarity
adf.test(xrp_open_ma, alternative = "stationary")
# Stop if p-value < 0.05

```

```{r }
xrpLog <- log(xrp_open_ma)
plot(xrpLog, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(xrpLog)
# Stop if p-value < 0.05
```

```{r }
xrpLogDiff <- diff(xrpLog)
plot(xrpLogDiff, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(xrpLogDiff)
# Stop if p-value < 0.05
```

### ACF and PACF of Ripple Open Moving Average

```{r }
op <- par(mfrow = c(1, 2))
acf(xrpLogDiff, main='ACF For XRP Differenced Series')
pacf(xrpLogDiff, main='PACF For XRP Differenced Series')
par(op)
```

### Arima of Open Price Based on Guess

```{r }
arimaFit <- arima(xrpLogDiff, order = c(2,0,0))
arimaFit            # check the coefficients
plot(arimaFit)


arimaFitFC <- forecast(xrpLogDiff, model = arimaFit, h = 10)
plot(xrpLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaFitFC$fitted, type="l", col = "red", lwd = 2)

plot(arimaFitFC)
```

```{r }
tsdisplay(residuals(arimaFit), lag.max = 80)
```

### Finding Optimal Arima of Ripple

```{r}
auto.arima(xrpLogDiff,                   # the dataset
           seasonal = FALSE,            # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

### Arima Optimal Plots

```{r }
arimaOpt <- arima(xrpLogDiff, order = c(1,0,0))
arimaOpt
plot(arimaOpt)

arimaOptFC <- forecast(xrpLogDiff, model = arimaOpt, h = 10)
plot(xrpLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaOptFC$fitted, type="l", col = "red", lwd = 2)
# Beyond the training dataset
plot(arimaOptFC)
```

### Ripple Seasonal? 

```{r}
auto.arima(xrpLogDiff,                   # the dataset
           seasonal = TRUE,             # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

No! Ripple is not seasonal and thus we just utilize the auto.arima above.

```{r}
# Check the performance/accuracy
tsdisplay(residuals(arimaOpt), lag.max = 80)
```

### Ripple Close Price Stationarity 

```{r}
xrp_close_ma <- ts(na.omit(xrp_latest_model$close_ma))

# Stationarity
adf.test(eth_close_ma, alternative = "stationary")
# Stop if p-value < 0.05
```

```{r}
xrpCloseLog <- log(xrp_close_ma)
plot(xrpCloseLog, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(xrpCloseLog)
# Stop if p-value < 0.05
```

```{r}
xrpCloseLogDiff <- diff(xrpCloseLog)
plot(xrpCloseLogDiff, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(xrpCloseLogDiff)
# Stop if p-value < 0.05
```

### Ripple Close Price's ACF and PACF

```{r }
op <- par(mfrow = c(1, 2))
acf(xrpCloseLogDiff, main='ACF For XRP Differenced Series')
pacf(xrpCloseLogDiff, main='PACF For XRP Differenced Series')
par(op)
```

### Ripple Close Price Arima Guess

```{r }
# Arima Based on guessing
arimaFit <- arima(xrpCloseLogDiff, order = c(4,0,2))
arimaFit            # check the coefficients
plot(arimaFit)

arimaFitFC <- forecast(xrpCloseLogDiff, model = arimaFit, h = 10)
plot(xrpCloseLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaFitFC$fitted, type="l", col = "red", lwd = 2)

plot(arimaFitFC)
```

```{r}
tsdisplay(residuals(arimaFit), lag.max = 80)
```

### Optimal Arima Close Price
```{r}
auto.arima(xrpCloseLogDiff,                   # the dataset
           seasonal = FALSE,            # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

### Ripple Close Forecast Plot

```{r}
arimaOpt <- arima(xrpCloseLogDiff, order = c(1,0,1))
arimaOpt
plot(arimaOpt)

# Forecast using optimal ARIMA
# Within the training dataset
arimaOptFC <- forecast(xrpCloseLogDiff, model = arimaOpt, h = 10)
plot(xrpCloseLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaOptFC$fitted, type="l", col = "red", lwd = 2)
# Beyond the training dataset
plot(arimaOptFC)
```

```{r}
tsdisplay(residuals(arimaOpt), lag.max = 80)
```

# Bitcoin Cash

```{r }
# plot open price
ggplot() + 
  geom_line(data = bch_sorted_final, aes(x = date, y = open, color="Open Price")) +
  geom_line(data = bch_sorted_final, aes(x = date, y = close, color="Close Price"))

summary(bch_sorted_final$open)
summary(bch_sorted_final$close)
```

### Moving Average of Bitcoin Cash [Weekly]

Since the price of cryptocurrencies are so volatile. I decided to use a rolling moving average to stabilize the data slightly.

```{r }
bch_latest_model <- bch_sorted_final[bch_sorted_final$date >= "2017-04-15" & bch_sorted_final$date <= "2018-04-15",]
bch_latest_model$open_ma = ma(bch_latest_model$open, order=7) # Weekly Open Moving Average

bch_latest_model$close_ma = ma(bch_latest_model$close, order=7) # Weekly Close Moving Average

ggplot() + 
  geom_line(data = bch_latest_model, aes(x = date, y = open, color="open prices")) +
  geom_line(data = bch_latest_model, aes(x = date, y = close, color="close prices")) +
  geom_line(data = bch_latest_model, aes(x = date, y = open_ma, color="weekly open ma")) + 
  geom_line(data = bch_latest_model, aes(x = date, y= close_ma, color="weekly close ma"))

```

### Creating Stationary Data for Open Moving Average 

```{r}
bch_open_ma <- ts(na.omit(bch_latest_model$open_ma))

# Stationarity
adf.test(bch_open_ma, alternative = "stationary")
# Stop if p-value < 0.05

```

```{r }
bchLog <- log(bch_open_ma)
plot(bchLog, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(bchLog)
# Stop if p-value < 0.05
```

```{r }
bchLogDiff <- diff(bchLog)
plot(bchLogDiff, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(bchLogDiff)
# Stop if p-value < 0.05
```

### ACF and PACF of Bitcoin Cash Open Moving Average

```{r }
op <- par(mfrow = c(1, 2))
acf(bchLogDiff, main='ACF For XRP Differenced Series')
pacf(bchLogDiff, main='PACF For XRP Differenced Series')
par(op)
```

### Arima of Open Price Based on Guess

```{r }
arimaFit <- arima(bchLogDiff, order = c(3,0,3))
arimaFit            # check the coefficients
plot(arimaFit)


arimaFitFC <- forecast(bchLogDiff, model = arimaFit, h = 10)
plot(bchLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaFitFC$fitted, type="l", col = "red", lwd = 2)

plot(arimaFitFC)
```

```{r }
tsdisplay(residuals(arimaFit), lag.max = 80)
```

### Finding Optimal Arima of Bitcoin Cash

```{r}
auto.arima(bchLogDiff,                   # the dataset
           seasonal = FALSE,            # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

### Arima Optimal Plots

```{r }
arimaOpt <- arima(bchLogDiff, order = c(2,0,3))
arimaOpt
plot(arimaOpt)

arimaOptFC <- forecast(bchLogDiff, model = arimaOpt, h = 10)
plot(bchLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaOptFC$fitted, type="l", col = "red", lwd = 2)
# Beyond the training dataset
plot(arimaOptFC)
```

### Bitcoin Cash Seasonal? 

```{r}
auto.arima(bchLogDiff,                   # the dataset
           seasonal = TRUE,             # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

No! Bitcoin Cash is not seasonal and thus we just utilize the auto.arima above.

```{r}
# Check the performance/accuracy
tsdisplay(residuals(arimaOpt), lag.max = 80)
```

### Bitcoin Cash Close Price Stationarity 

```{r}
bch_close_ma <- ts(na.omit(bch_latest_model$close_ma))

# Stationarity
adf.test(bch_close_ma, alternative = "stationary")
# Stop if p-value < 0.05
```

```{r}
bchCloseLog <- log(bch_close_ma)
plot(bchCloseLog, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(bchCloseLog)
# Stop if p-value < 0.05
```

```{r}
bchCloseLogDiff <- diff(bchCloseLog)
plot(bchCloseLogDiff, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(bchCloseLogDiff)
# Stop if p-value < 0.05
```

### Bitcoin Cash Close Price's ACF and PACF

```{r }
op <- par(mfrow = c(1, 2))
acf(bchCloseLogDiff, main='ACF For BCH Differenced Series')
pacf(bchCloseLogDiff, main='PACF For BCH Differenced Series')
par(op)
```

### Bitcoin Cash Close Price Arima Guess

```{r }
# Arima Based on guessing
arimaFit <- arima(bchCloseLogDiff, order = c(4,0,2))
arimaFit            # check the coefficients
plot(arimaFit)

arimaFitFC <- forecast(bchCloseLogDiff, model = arimaFit, h = 10)
plot(bchCloseLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaFitFC$fitted, type="l", col = "red", lwd = 2)

plot(arimaFitFC)
```

```{r}
tsdisplay(residuals(arimaFit), lag.max = 80)
```

### Optimal Arima Close Price
```{r}
auto.arima(bchCloseLogDiff,                   # the dataset
           seasonal = FALSE,            # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

### Bitcoin Cash Close Forecast Plot

```{r}
arimaOpt <- arima(bchCloseLogDiff, order = c(1,0,1))
arimaOpt
plot(arimaOpt)

# Forecast using optimal ARIMA
# Within the training dataset
arimaOptFC <- forecast(bchCloseLogDiff, model = arimaOpt, h = 10)
plot(bchCloseLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaOptFC$fitted, type="l", col = "red", lwd = 2)
# Beyond the training dataset
plot(arimaOptFC)
```

```{r}
tsdisplay(residuals(arimaOpt), lag.max = 80)
```

# Litecoin

```{r }
# plot open price
ggplot() + 
  geom_line(data = ltc_sorted_final, aes(x = date, y = open, color="Open Price")) +
  geom_line(data = ltc_sorted_final, aes(x = date, y = close, color="Close Price"))

summary(ltc_sorted_final$open)
summary(ltc_sorted_final$close)
```

### Moving Average of Litecoin [Weekly]

Since the price of cryptocurrencies are so volatile. I decided to use a rolling moving average to stabilize the data slightly.

```{r }
ltc_latest_model <- ltc_sorted_final[ltc_sorted_final$date >= "2017-04-15" & ltc_sorted_final$date <= "2018-04-15",]
ltc_latest_model$open_ma = ma(ltc_latest_model$open, order=7) # Weekly Open Moving Average

ltc_latest_model$close_ma = ma(ltc_latest_model$close, order=7) # Weekly Close Moving Average

ggplot() + 
  geom_line(data = ltc_latest_model, aes(x = date, y = open, color="open prices")) +
  geom_line(data = ltc_latest_model, aes(x = date, y = close, color="close prices")) +
  geom_line(data = ltc_latest_model, aes(x = date, y = open_ma, color="weekly open ma")) + 
  geom_line(data = ltc_latest_model, aes(x = date, y= close_ma, color="weekly close ma"))

```

### Creating Stationary Data for Open Moving Average 

```{r}
ltc_open_ma <- ts(na.omit(ltc_latest_model$open_ma))

# Stationarity
adf.test(ltc_open_ma, alternative = "stationary")
# Stop if p-value < 0.05

```

```{r }
ltcLog <- log(ltc_open_ma)
plot(ltcLog, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(ltcLog)
# Stop if p-value < 0.05
```

```{r }
ltcLogDiff <- diff(ltcLog)
plot(ltcLogDiff, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(ltcLogDiff)
# Stop if p-value < 0.05
```

### ACF and PACF of LiteCoin Open Moving Average

```{r }
op <- par(mfrow = c(1, 2))
acf(ltcLogDiff, main='ACF For BCH Differenced Series')
pacf(ltcLogDiff, main='PACF For BCH Differenced Series')
par(op)
```

### Arima of Open Price Based on Guess

```{r }
arimaFit <- arima(ltcLogDiff, order = c(3,0,2))
arimaFit            # check the coefficients
plot(arimaFit)


arimaFitFC <- forecast(ltcLogDiff, model = arimaFit, h = 10)
plot(ltcLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaFitFC$fitted, type="l", col = "red", lwd = 2)

plot(arimaFitFC)
```

```{r }
tsdisplay(residuals(arimaFit), lag.max = 80)
```

### Finding Optimal Arima of Litecoin

```{r}
auto.arima(ltcLogDiff,                   # the dataset
           seasonal = FALSE,            # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

### Arima Optimal Plots

```{r }
arimaOpt <- arima(ltcLogDiff, order = c(2,0,1))
arimaOpt
plot(arimaOpt)

arimaOptFC <- forecast(ltcLogDiff, model = arimaOpt, h = 10)
plot(ltcLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaOptFC$fitted, type="l", col = "red", lwd = 2)
# Beyond the training dataset
plot(arimaOptFC)
```

### LiteCoin Seasonal? 

```{r}
auto.arima(ltcLogDiff,                   # the dataset
           seasonal = TRUE,             # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

No! Litecoin is not seasonal and thus we just utilize the auto.arima above.

```{r}
# Check the performance/accuracy
tsdisplay(residuals(arimaOpt), lag.max = 80)
```

### LiteCoin Close Price Stationarity 

```{r}
ltc_close_ma <- ts(na.omit(ltc_latest_model$close_ma))

# Stationarity
adf.test(ltc_close_ma, alternative = "stationary")
# Stop if p-value < 0.05
```

```{r}
ltcCloseLog <- log(ltc_close_ma)
plot(ltcCloseLog, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(ltcCloseLog)
# Stop if p-value < 0.05
```

```{r}
ltcCloseLogDiff <- diff(ltcCloseLog)
plot(ltcCloseLogDiff, type="l", col = "blue", 
     lwd = 3, xlab = "", ylab = "")
adf.test(ltcCloseLogDiff)
# Stop if p-value < 0.05
```

### Litecoin Close Price's ACF and PACF

```{r }
op <- par(mfrow = c(1, 2))
acf(ltcCloseLogDiff, main='ACF For XRP Differenced Series')
pacf(ltcCloseLogDiff, main='PACF For XRP Differenced Series')
par(op)
```

### Litecoin Close Price Arima Guess

```{r }
# Arima Based on guessing
arimaFit <- arima(ltcCloseLogDiff, order = c(4,0,2))
arimaFit            # check the coefficients
plot(arimaFit)

arimaFitFC <- forecast(ltcCloseLogDiff, model = arimaFit, h = 10)
plot(ltcCloseLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaFitFC$fitted, type="l", col = "red", lwd = 2)

plot(arimaFitFC)
```

```{r}
tsdisplay(residuals(arimaFit), lag.max = 80)
```

### Optimal Arima Close Price
```{r}
auto.arima(ltcCloseLogDiff,                   # the dataset
           seasonal = FALSE,            # seasonality
           stationary = TRUE,           # stationarity
           max.p = 5, max.q = 5,        # range of p,q (non-seasonal)
           max.P = 5, max.Q = 5,        # range of P,Q (seasonal)
           max.d = 2, max.D = 2,        # range of d,D (differences)
           start.p = 1, start.q = 1,    # start for stepwise search
           start.P = 1, start.Q = 1,    # start for stepwise search 
           ic = "aic",                  # criteria to compare
           stepwise = TRUE,             # stepwise selection (faster)
           trace = TRUE)                # all ARIMA models reported
```

### LiteCoin Close Forecast Plot

```{r}
arimaOpt <- arima(ltcCloseLogDiff, order = c(1,0,1))
arimaOpt
plot(arimaOpt)

# Forecast using optimal ARIMA
# Within the training dataset
arimaOptFC <- forecast(ltcCloseLogDiff, model = arimaOpt, h = 10)
plot(ltcCloseLogDiff, type="l", col = "lightblue", 
     lwd = 3, xlab = "", ylab = "")
points(arimaOptFC$fitted, type="l", col = "red", lwd = 2)
# Beyond the training dataset
plot(arimaOptFC)
```

```{r}
tsdisplay(residuals(arimaOpt), lag.max = 80)
```
